Cover Your Basis: Comprehensive Data-Driven Characterization of the Binary Black Hole Population

### 2022-10-11 src/tex/results.tex
remove sentance about fitting chieff with splines

We model the spin magnitude distributions as IID B-Spline distributions as a flexible check to the assumptions of the \textsc{Default} model. 
Figure \ref{fig:iid_spinmag_dist} shows the inferred spin magnitude distribution with the B-Spline model, compared with the \textsc{Default} model. 
The B-Spline model finds consistent agreement with the beta distribution, peaking near $a\sim0.2$, with 90\% of BBH spins below 
\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][a_90percentile]}$} at 90\% credibility. The B-Spline model has no enforcement for vanishing support at the extremal values like the Beta distribution, 
allowing it to probe the zero-spin question. We find broad support, with large variance, for non-zero probabilities at $a_i=0$, but cannot confidently determine the presence of 
a significant non-spinning subpopulation, corroborating similar recent conclusions \citep{BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}. 
We now repeat the same analysis as above but with independent B-Spline distributions for each spin magnitude component. In figure \ref{fig:ind_spinmag_dist} 
we show the inferred primary (orange), and secondary (olive) spin magnitude distributions inferred when relaxing the IID assumption. We find that the primary spin magnitude distribution 
peaking slightly higher at $a\sim0.25$ than the IID B-Spline model, but with similar support at near vanishing spins. The secondary spin magnitude distribution is more uncertain due to the 
higher measurement uncertainty in inferring the secondary spins of BBH systems. The secondary distribution also peaks at smaller spin magnitudes of $a\sim0.15$, showing considerably larger 
PPD and confidence band at $a\sim0$ than the primary distribution or B-Spline IID spin magnitude distribution in figure \ref{fig:iid_spinmag_dist}. While the distributions are still broadly 
consistent with each other this could also be signs pointing towards component spin magnitude distributions uniquely distributed.

@@

While the component spin magnitude and tilts are more easily interpretable in context of the astrophysics, they are parameters that 
are not measured well during individual event parameter estimation. The best measured spin parameter, that enters the waveform at XXPN, is 
called the effective spin defined as: $\chi_\mathrm{eff} = \frac{a_1\cos{\theta_1} + qa_2\cos{\theta_2}}{1+q}$. As an alternative check 
on the spin distributions we also try modeling an B-Spline distribution of effective spins instead of component spin magnitude and tilts 
as done above. Figure \ref{fig:chieff_dist} shows the inferred effective spin distribution with the B-Spline models (red and purple) 
along with results on the \textsc{Default} and \textsc{Gaussian} models. We find considerable agreement 
among the effective spin distributions, but the more flexible B-Spline models in component spins, more closely resemble the \textsc{Gaussian} 
that directly models $\chi_\mathrm{eff}$, than the \textsc{Default} model. This could be signs that the population fit is driven by $\chi_\mathrm{eff}$ 
more than the component spin dimensions. The B-Spline model finds very similar shapes to the other models, with a peak centered at 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][iid][PeakChiEff]}$}, compared to 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][default][PeakChiEff]}$} with the \textsc{Default} model and 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][gaussian][PeakChiEff]}$} with the \textsc{Gaussian} $\chi_\mathrm{eff}$ model. 
As for the misaligned spin question, we calculate the fraction of systems with effective spins that are misaligned (i.e. $\chi_\mathrm{eff}<0$) and 
find for the B-Spline model $f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][iid][FracBelow0]}$}, compared to 
$f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][default][FracBelow0]}$} with the \textsc{Default} model and 
$f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][gaussian][FracBelow0]}$} with the Gaussian $\chi_\mathrm{eff}$ model.

>>>>

We model the spin magnitude distributions as IID B-Spline distributions as a flexible check to the assumptions of the \textsc{Default} model. 
Figure \ref{fig:iid_spinmag_dist} shows the inferred spin magnitude distribution with the B-Spline model, compared with the \textsc{Default} model. 
The B-Spline model finds consistent agreement with the beta distribution, peaking near $a\sim0.2$, with 90\% of BBH spins below 
\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][a_90percentile]}$} at 90\% credibility. The B-Spline model has no enforcement for vanishing support at the extremal values like the Beta distribution, 
allowing it to probe the zero-spin question. We find broad support, with large variance, for non-zero probabilities at $a_i=0$, but cannot confidently determine the presence of 
a significant non-spinning subpopulation, corroborating similar recent conclusions \citep{BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}. 
We now repeat the same analysis as above but with independent B-Spline distributions for each spin magnitude component. In figure \ref{fig:ind_spinmag_dist} 
we show the inferred primary (orange), and secondary (olive) spin magnitude distributions inferred when relaxing the IID assumption. We find that the primary spin magnitude distribution 
peaking slightly higher at $a\sim0.25$ than the IID B-Spline model, but with similar support at near vanishing spins. The secondary spin magnitude distribution is more uncertain due to the 
higher measurement uncertainty in inferring the secondary spins of BBH systems. The secondary distribution also peaks at smaller spin magnitudes of $a\sim0.15$, showing considerably larger 
PPD and confidence band at $a\sim0$ than the primary distribution or B-Spline IID spin magnitude distribution in figure \ref{fig:iid_spinmag_dist}. While the distributions are still broadly 
consistent with each other this could also be signs pointing towards component spin magnitude distributions being uniquely distributed.

@@

While the component spin magnitude and tilts are more easily interpretable in context of the astrophysics, they are parameters that 
are not measured well during individual event parameter estimation. The best measured spin parameter, that enters the waveform at XXPN, is 
called the effective spin defined as: $\chi_\mathrm{eff} = \frac{a_1\cos{\theta_1} + qa_2\cos{\theta_2}}{1+q}$. Figure \ref{fig:chieff_dist} 
shows the inferred effective spin distribution with the two versions of our B-Spline models (red and purple), 
along with results on the \textsc{Default} and \textsc{Gaussian} models from \citet{o3b_astro_dist}. We find considerable agreement 
among the effective spin distributions, but the more flexible B-Spline models in component spins, more closely resemble the \textsc{Gaussian} 
that directly models $\chi_\mathrm{eff}$, than the \textsc{Default} model. This could be signs that the population fit is driven by $\chi_\mathrm{eff}$ 
more than the component spin dimensions. The B-Spline model finds very similar shapes to the other models, with a peak centered at 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][iid][PeakChiEff]}$}, compared to 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][default][PeakChiEff]}$} with the \textsc{Default} model and 
$\chi_\mathrm{eff}=$\result{$\CIPlusMinus{\macros[ChiEffective][gaussian][PeakChiEff]}$} with the \textsc{Gaussian} $\chi_\mathrm{eff}$ model. 
As for the misaligned spin question, we calculate the fraction of systems with effective spins that are misaligned (i.e. $\chi_\mathrm{eff}<0$) and 
find for the B-Spline model $f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][iid][FracBelow0]}$}, compared to 
$f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][default][FracBelow0]}$} with the \textsc{Default} model and 
$f_{\chi_\mathrm{eff}<0}=$\result{$\CIPlusMinus{\macros[ChiEffective][gaussian][FracBelow0]}$} with the Gaussian $\chi_\mathrm{eff}$ model.
